### 2022-09-06 main/chapter_searches.tex
write on p-value ratios

Fisher illustrates the use of $p$-values in the second example
of~\cite{fisher1921probable}, which models some data with a parameter $r$ and
compares two hypotheses
$r=0.3$ and
$r=0.95$, finding $p$-values
$P(r=0.3)=0.142$ and
$P(r=0.95)=0.00014$.
``Thus'', Fisher concludes, ``the value $0.95$\ldots
is roughly $1000$ times less likely than $0.30$''
(``when measured on the $r$ scale'').
By writing `likely', Fisher is carefully not referencing the probability
of parameters, but their likelihoods in our definition.
But this factor of $1000$ is a $p$-value ratio, not a likelihood ratio,
so how can this make sense?
Plausibly, it is a data reduction; comparing
Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
shows that the $p$-value is a likelihood (only) if the result is reduced from
$A: t =    \tilde{t}$ to
$B: t \geq \tilde{t}$.
And since this reduction is technically true --- $A$ implies $B$ --- the
likelihood ratio on $B$ is logically valid and not \emph{necessarily}
misleading.

But $B: t \geq \tilde{t}$ is a logical statement (true for all test statistics
greater than or equal to the observed value) and not a randomly occurring
event, so its assignment frequentist probability is deeply dubious.
Much more importantly, the reduction to a one-sided inequality has no tangible
benefit; it reduces the information content of the reported data, but does not
reduce the costs of reporting or interpreting those data.
Both $A$ and $B$ are printed in the same number of characters, and $p$-values
can be painfully expensive to evaluate.
Furthermore, it can actively harm the result by diluting the likelihood ratio,
since if
\begin{equation}
\phi = \log \frac{L(h_1)}{L(h_2)}
\to
\log \frac{
L(h_1) + \prob{t > \tilde{t}}{h_1}
}
{
L(h_2) + \prob{t > \tilde{t}}{h_2}
}
\end{equation}
then large tail probabilities will only squash the $\log$ ratio toward zero.

>>>>

Fisher illustrates the use of $p$-values in the second example
of~\cite{fisher1921probable}, which models some data with a parameter $r$ and
compares two hypotheses
$r=0.3$ and
$r=0.95$, finding $p$-values
$P(r=0.3)=0.142$ and
$P(r=0.95)=0.00014$.
``Thus'', Fisher concludes, ``the value $0.95$\ldots
is roughly $1000$ times less likely than $0.30$''
(``when measured on the $r$ scale'').
By writing `likely', Fisher is carefully not referencing the probability
of parameters, but their likelihoods in our definition.
But this factor of $1000$ is a $p$-value ratio, not a likelihood ratio,
so how can this make sense?
Plausibly, it is a data reduction; comparing
Equations~\ref{eqn:searches_pvalue} and~\ref{eqn:searches_likelihood}
shows that the $p$-value is a likelihood (only) if the result is reduced from
$A: t =    \tilde{t}$ to
$B: t \geq \tilde{t}$.
And since this reduction is technically true --- $A$ implies $B$ --- the
likelihood ratio on $B$ is logically valid and not \emph{necessarily}
misleading if clearly stated.

But $B: t \geq \tilde{t}$ is a logical statement (true for all test statistics
greater than or equal to the observed value) and not a randomly occurring
event, so its assignment frequentist probability is deeply dubious.
Much more importantly, the reduction to a one-sided inequality has no tangible
benefit; it reduces the information content of the reported data, but does not
reduce the costs of reporting or interpreting those data.
Both $A$ and $B$ are printed in the same number of characters, and $p$-values
can be painfully expensive to evaluate.
Furthermore, it actively harms any result with small a test statistic by
diluting the likelihood ratio, since if
\begin{equation}
\log \frac{L(h_1)}{L(h_2)}
\to
\log \frac{
L(h_1) + \prob{t > \tilde{t}}{h_1}
}
{
L(h_2) + \prob{t > \tilde{t}}{h_2}
}
\end{equation}
then large tail probabilities will only squash the $\log$ likelihood ratio
toward zero.
